---
title: "Modeling"
author: "Philipp Bayer"
date: "18 March 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# First, setup

```{r}
library(tidyverse)
library(cowplot)
library(minpack.lm)
library(broom)
library(car)
library(patchwork)

options('scipen'=100)
```


## Oleracea model

Here I load the oleracea data and plot it for oleracea only
```{r oleracea_model}
oleracea_df <- read_tsv('./Oleracea_PAV_nice_names_intersection_of_individuals.filtered.onlyInds.86Permuts.txt')

# approximate the para0p
lm(logit(pangenome/60000) ~ Combinations, oleracea_df)

para0p <- c(A=60000,B=-3.36,C=0.002)
```
OK we got the initaliser

```{r oleracea_training}
oleracea_summary_fitp <- nlsLM(pangenome~A*Combinations^B+C, data=oleracea_df, 
                      start=para0p, trace=F, control=list(maxiter=200, maxfev=1500))

summary(oleracea_summary_fitp)
tidy(oleracea_summary_fitp)
```
Now come the core genes!

```{r oleracea_core}
para0c <- c(B=1)
oleracea_fitc <- nls(core~cbind(1, exp(B*Combinations)),oleracea_df, start=para0c,             trace=F, alg='plinear')
tidy(oleracea_fitc)
```

```{r oleracea_plt}
oleracea_fit_data_p <- augment(oleracea_summary_fitp)
oleracea_fit_data_c <- augment(oleracea_fitc)

oleracea_p <-ggplot() + 
  geom_point(data=oleracea_df, aes(x=Combinations, y=pangenome), alpha=0.1) + 
  geom_point(data=oleracea_df, aes(x=Combinations, y=core), alpha=0.1) + 
  geom_line(data=oleracea_fit_data_p, aes(x=Combinations, y=.fitted, colour='red')) + 
  geom_line(data=oleracea_fit_data_c, aes(x=Combinations, y=.fitted, colour='red')) + 
  #xlim(c(0,50)) +
  ylab('Number of pangenome genes') +
  xlab('Combinations of individuals') +
  theme(legend.position = "none") +
  theme(aspect.ratio = 1) +
  theme_half_open() +
  theme_minimal_hgrid() +
  theme(legend.position="none")+
  ylim(c(39000, 60000))
oleracea_p
```

## Rapa model

```{r rapa_model}
rapa_df <- read_tsv('./Rapa_PAV_nice_names_intersection_of_individuals.filtered.onlyInds.71Permuts.txt')
head(rapa_df)

# approximate the para0p

lm(logit(pangenome/59000) ~ Combinations, rapa_df)

para0p <- c(A=59000,B=-4.59,C=0.00022)
```
OK we got the initaliser

```{r rapa_training}
rapa_summary_fitp <- nlsLM(pangenome~A*Combinations^B+C, data=rapa_df, 
                      start=para0p, trace=F, control=list(maxiter=200, maxfev=1500))

summary(rapa_summary_fitp)
tidy(rapa_summary_fitp)
```
Now come the core genes!

```{r rapa_core}

para0c <- c(B=1)
rapa_fitc <- nls(core~cbind(1, exp(B*Combinations)),rapa_df, start=para0c, 
            trace=F, alg='plinear')
tidy(rapa_fitc)
```

```{r rapa_plot}
rapa_fit_data_p <- augment(rapa_summary_fitp)
rapa_fit_data_c <- augment(rapa_fitc)

rapa_p <- ggplot() + 
  geom_point(data=rapa_df, aes(x=Combinations, y=pangenome), alpha=0.1) +
  geom_point(data=rapa_df, aes(x=Combinations, y=core), alpha=0.1) +
  geom_line(data=rapa_fit_data_p, aes(x=Combinations, y=.fitted, colour='red')) +
  geom_line(data=rapa_fit_data_c, aes(x=Combinations, y=.fitted, colour='red')) +
  ylab('Number of pangenome genes') +
  xlab('Combinations of individuals') +
  theme(legend.position = "none") +
  theme(aspect.ratio = 1) +
  theme_half_open() +
  theme_minimal_hgrid() +
  theme(legend.position="none") +
  ylim(c(39000, 60000))
rapa_p
```



## Napus model

```{r napus_model}
napus_df <- read_tsv('./NewPAV_Table_nice_names_intersection_of_individuals.filtered.onlyInds.72Permuts.txt')

# approximate the para0p

lm(logit(pangenome/(max(napus_df$pangenome)+2000)) ~ Combinations, napus_df)

para0p <- c(A=max(napus_df$pangenome)+1000,B=-3.754,C=0.004609)

```
OK we got the initaliser

```{r napus_training}
napus_summary_fitp <- nlsLM(pangenome~A*Combinations^B+C, data=napus_df, 
                      start=para0p, trace=F, control=list(maxiter=200, maxfev=1500))

summary(napus_summary_fitp)
tidy(napus_summary_fitp)
```

Now come the core genes!

```{r napus_core}

para0c <- c(B=1)
napus_fitc <- nls(core~cbind(1, exp(B*Combinations)),napus_df, start=para0c, 
            trace=F, alg='plinear')
tidy(napus_fitc)

```


```{r napus_plot}
napus_fit_data_p <- augment(napus_summary_fitp)
napus_fit_data_c <- augment(napus_fitc)

napus_p <- ggplot() + 
  geom_point(data=napus_df, aes(x=Combinations, y=pangenome), alpha=0.1) + 
  geom_point(data=napus_df, aes(x=Combinations, y=core), alpha=0.1) + 
  geom_line(data=napus_fit_data_p, aes(x=Combinations, y=.fitted, colour='red')) +
  geom_line(data=napus_fit_data_c, aes(x=Combinations, y=.fitted, colour='red')) +
  ylab('Number of pangenome genes') +
  xlab('Combinations of individuals') +
  theme(legend.position = "none") +
  theme(aspect.ratio=1) +
  theme_half_open() +
  theme_minimal_hgrid() +
  theme(legend.position="none") +
  ylim(c(65000, 110000))

napus_p
```


## Napus no synths model

```{r}
napus_df <- read_tsv('./NewPAV_Table_nice_names_intersection_of_individuals.filtered.onlyInds.NoSynths.54Inds.txt')
tail(napus_df)

# approximate the para0p
max(napus_df$pangenome)
lm(logit(pangenome/105000) ~ Combinations, napus_df)

```
OK we got the initaliser

```{r }
para0p <- c(A=110000, B=-4.57,C=0.00022)


napus_summary_fitp <- nlsLM(pangenome~A*Combinations^B+C, data=napus_df, 
                      start=para0p, trace=F, control=list(maxiter=200, maxfev=1500))
summary(napus_summary_fitp)
tidy(napus_summary_fitp)
```

Now come the core genes!

```{r }
para0c <- c(B=1)
napus_fitc <- nls(core~cbind(1, exp(B*Combinations)),napus_df, start=para0c, 
            trace=F, alg='plinear')
tidy(napus_fitc)

```


```{r }
napus_fit_data_p <- augment(napus_summary_fitp)
napus_fit_data_c <- augment(napus_fitc)

napus_nonsyn_p <- ggplot() + 
  geom_point(data=napus_df, aes(x=Combinations, y=pangenome), alpha=0.1) + 
  geom_point(data=napus_df, aes(x=Combinations, y=core), alpha=0.1) + 
  geom_line(data=napus_fit_data_p, aes(x=Combinations, y=.fitted, colour='red')) +
  geom_line(data=napus_fit_data_c, aes(x=Combinations, y=.fitted, colour='red')) +
  ylab('Number of pangenome genes') +
  xlab('Combinations of individuals') +
  theme(legend.position = "none") +
  theme(aspect.ratio=1) +
  theme_half_open() +
  theme_minimal_hgrid() +
  theme(legend.position="none") +
  ylim(c(65000, 110000))

napus_nonsyn_p
```


# Patchworking!

```{r patches}
grid <- plot_grid(oleracea_p, rapa_p, napus_p, napus_nonsyn_p, labels='AUTO')
grid
save_plot(filename= 'Grid.png', plot = grid, base_height = 8)
```

# All in one!

```{r all_in_one}
ggplot() + 
  geom_point(data=napus_df, aes(x=Combinations, y=pangenome), alpha=0.1) + 
  geom_line(data=napus_fit_data_p, aes(x=Combinations, y=.fitted, colour='red')) +
  geom_point(data=rapa_df, aes(x=Combinations, y=pangenome), alpha=0.1) +
    geom_line(data=rapa_fit_data_p, aes(x=Combinations, y=.fitted, colour='red')) +
  geom_point(data=oleracea_df, aes(x=Combinations, y=pangenome), alpha=0.1) +
    geom_line(data=oleracea_fit_data_p, aes(x=Combinations, y=.fitted, colour='red')) +
  ylab('Number of pangenome genes') +
  xlab('Combinations of individuals') +
  theme(legend.position = "none") +
  theme(aspect.ratio=1) +
  theme_half_open() +
  theme_minimal_hgrid() +
  theme(legend.position="none") +
  xlim(c(0, 50))

```

yeah let's not use that :P



