```{r setup}
library(tidyverse)
library(cowplot)
library(patchwork)
library(ggtext)
```

OK, let's first load the data.

# Loading

```{r}
napus_X <- read_tsv('./../9-PAVModeling/Napus_Feature_Table.Filtered.csv')
# need to delete the outcome columns to make the fit with shap-values work
napus_X$`Variable in synth` <- NULL
napus_X$Name <- NULL
napus_X$Variable <- NULL
head(napus_X)
```
```{r}
# this file was made in the jupyter model notebook
napus_shap <- read_csv('./napus_shap_values.csv', col_names = F)
```
```{r}
head(napus_shap)
# each column is one phenotype, each row is one gene, same dimensions as the model input table
```

So napus_shap$X1 corresponds with napus_X chrA01, I guess?
```{r}
names(napus_shap) <- names(napus_X)
```

Let's make a test plot
```{r}
plot(napus_X$`Position on chromosome`, napus_shap$`Position on chromosome`)
```

Nice, that's perfect. 

# Napus plot

Now I want to remake my plots - first,  pull out A01 all rows where A01 == 1 in napus_X

```{r}
napus_x_a01 <- napus_X %>% filter(chrA01 == 1)
napus_shap_a01 <- napus_shap %>% filter(napus_X$chrA01 == 1)
head(napus_shap_a01)
head(napus_x_a01)
nap_a01_df <- tibble(`Position on chromosome` = napus_x_a01$`Position on chromosome`,
             `SHAP value` = napus_shap_a01$`Position on chromosome`)


nap_a01_m <- lm(`SHAP value` ~ `Position on chromosome`, nap_a01_df)

#eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
#     list(a = format(unname(coef(m)[1]), digits = 2),
#          b = format(unname(coef(m)[2]), digits = 2),
#         r2 = format(summary(m)$r.squared, digits = 3)))
#as.character(as.expression(eq))

napus_a01_p <-
  ggplot(data = nap_a01_df, aes(x = `Position on chromosome`, y = `SHAP value`)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = 'lm', formula = y ~ x) +
  #geom_text(x = 0.2, y=1.4, label = as.character(as.expression(eq)), parse=T) +
  theme_minimal_hgrid() +
  ylim(c(-0.65, 1.7)) +
  xlab('Position on *B. napus* A01') +
  theme(axis.title.x = element_markdown())

napus_a01_p
```

Good, now let's make the C01 plot too
```{r}
napus_x_c01 <- napus_X %>% filter(chrC01 == 1)
napus_shap_c01 <- napus_shap %>% filter(napus_X$chrC01 == 1)
head(napus_shap_c01)
head(napus_x_c01)
nap_c01_df <- tibble(`Position on chromosome` = napus_x_c01$`Position on chromosome`,
             `SHAP value` = napus_shap_c01$`Position on chromosome`)


nap_c01_m <- lm(`SHAP value` ~ `Position on chromosome`, nap_c01_df)

#eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
#     list(a = format(unname(coef(m)[1]), digits = 2),
#          b = format(unname(coef(m)[2]), digits = 2),
#         r2 = format(summary(m)$r.squared, digits = 3)))
#as.character(as.expression(eq))

napus_c01_p <-
  ggplot(data = nap_c01_df, aes(x = `Position on chromosome`, y = `SHAP value`)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = 'lm', formula = y ~ x) +
  #geom_text(x = 0.2, y=0.9, label = as.character(as.expression(eq)), parse=T) +
  theme_minimal_hgrid() +
  ylim(c(-0.65, 1.7)) +
  xlab('Position on *B. napus* C01') +
  theme(axis.title.x = element_markdown())

napus_c01_p
```

# Rapa plot

OK now comes the plot for Rapa A01

```{r}
rapa_X <- read_tsv('./../9-PAVModeling/Rapa_Feature_Table.Filtered.csv')
# need to delete the outcome columns to make the fit with shap-values work

rapa_X$Name <- NULL
rapa_X$Variable <- NULL
head(rapa_X)
```

```{r}
rapa_shap <- read_csv('./rapa_shap_values.csv', col_names = F)
names(rapa_shap) <- names(rapa_X)
```

```{r}
head(rapa_shap)
```

```{r}
rapa_x_a01 <- rapa_X %>% filter(A01 == 1)

rapa_shap_a01 <- rapa_shap %>% filter(rapa_X$A01 == 1)
head(rapa_shap_a01)
head(rapa_x_a01)

rap_df <- tibble(`Position on chromosome` = rapa_x_a01$`Position on chromosome`,
             `SHAP value` = rapa_shap_a01$`Position on chromosome`)


rap_m <- lm(`SHAP value` ~ `Position on chromosome`, rap_df)

#eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
#     list(a = format(unname(coef(m)[1]), digits = 2),
#          b = format(unname(coef(m)[2]), digits = 2),
#         r2 = format(summary(m)$r.squared, digits = 3)))
#as.character(as.expression(eq))

rapa_a01_p <-
  ggplot(data = rap_df, aes(x = `Position on chromosome`, y = `SHAP value`)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = 'lm') +
  #geom_text(x = 0.2, y=-0.5, label = as.character(as.expression(eq)), parse=T) +
  theme_minimal_hgrid() +
  ylim(c(-0.65, 1.7)) +
  xlab('Position on *B. rapa* A01') +
  theme(axis.title.x = element_markdown())

rapa_a01_p
```



# Oleracea plot

OK now comes the plot for Oleracea C1

```{r}
oleracea_X <- read_tsv('./../9-PAVModeling/Oleracea_Feature_Table.Filtered.csv')
# need to delete the outcome columns to make the fit with shap-values work

oleracea_X$Name <- NULL
oleracea_X$Variable <- NULL
head(oleracea_X)
```

```{r}
oleracea_shap <- read_csv('./oleracea_shap_values.csv', col_names = F)
names(oleracea_shap) <- names(oleracea_X)
```

```{r}
head(oleracea_shap)
```

```{r}
oleracea_x_c01 <- oleracea_X %>% filter(C1 == 1)

oleracea_shap_c01 <- oleracea_shap %>% filter(oleracea_X$C1 == 1)
head(oleracea_shap_c01)
head(oleracea_x_c01)

oler_df <-
  tibble(
  `Position on chromosome` = oleracea_x_c01$`Position on chromosome`,
  `SHAP value` = oleracea_shap_c01$`Position on chromosome`
  )
  
oler_m <- lm(`SHAP value` ~ `Position on chromosome`, oler_df)
  
#eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
#     list(a = format(unname(coef(m)[1]), digits = 2),
#          b = format(unname(coef(m)[2]), digits = 2),
#         r2 = format(summary(m)$r.squared, digits = 3)))
#as.character(as.expression(eq))

oleracea_c01_p <-
  ggplot(data = oler_df, aes(x = `Position on chromosome`, y = `SHAP value`)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = 'lm') +
  #geom_text(x = 0.35, y=0.3, label = as.character(as.expression(eq)), parse=T) +
  theme_minimal_hgrid() +
  ylim(c(-0.65, 1.7)) +
  xlab('Position on *B. oleracea* C01') +
  theme(axis.title.x = element_markdown())
oleracea_c01_p
```


# Now make a grid
```{r}
grid <-
  plot_grid(
  rapa_a01_p,
  oleracea_c01_p,
  napus_a01_p,
  napus_c01_p,
  align = 'vh',
  labels = c('A', 'B', 'C', 'D')
  )

grid

```

```{r}
ggsave('Napus_oleracea_rapa_grid.png', plot=grid, scale = 2)
```
# Merging all chromosomes into lines

Now let's try to merge those
```{r}

chrom_shap <- napus_shap$`Position on chromosome`
chrom_x <- napus_X %>% select(starts_with('chr'), `Position on chromosome`)
chrom_x <- chrom_x %>% pivot_longer(-`Position on chromosome`) %>% filter(value == 1)
chrom_x$value <- NULL
chrom_x$shap <- chrom_shap
head(chrom_x)
```
```{r}
paste(max(chrom_x$shap), min(chrom_x$shap))
```

```{r}
napus_lines <-
  ggplot(data = chrom_x, aes(x = `Position on chromosome`, y = shap, color = name)) +
  geom_line(alpha = 0.25) +
  theme_minimal_hgrid() +
  ylim(c(-0.65, 1.6)) +
  guides(color = FALSE)
  napus_lines
```

```{r}
chrom_rapa_shap <- rapa_shap$`Position on chromosome`
chrom_rapa_x <- rapa_X %>% select(starts_with('A'), `Position on chromosome`)
chrom_rapa_x <- chrom_rapa_x %>% pivot_longer(-`Position on chromosome`) %>% filter(value == 1)
chrom_rapa_x$value <- NULL
chrom_rapa_x$shap <- chrom_rapa_shap
head(chrom_rapa_x)
```
```{r}
paste(max(chrom_rapa_x$shap), min(chrom_rapa_x$shap))
```
```{r}
rapa_lines <-
  ggplot(data = chrom_rapa_x, aes(x = `Position on chromosome`, y = shap, color = name)) +
  geom_line(alpha = 0.25) +
  theme_minimal_hgrid() +
  ylim(c(-0.65, 1.6)) +
  guides(color = FALSE)
```

```{r}
chrom_oleracea_shap <- oleracea_shap$`Position on chromosome`
chrom_oleracea_x <- oleracea_X %>% select(starts_with('C'), `Position on chromosome`)
chrom_oleracea_x <- chrom_oleracea_x %>% pivot_longer(-`Position on chromosome`) %>% filter(value == 1)
chrom_oleracea_x$value <- NULL
chrom_oleracea_x$shap <- chrom_oleracea_shap
head(chrom_oleracea_x)
```

```{r}
oler_lines <- ggplot(data = chrom_oleracea_x, aes(x=`Position on chromosome`, y=shap, color = name)) + 
  geom_line(alpha=0.25) +
  theme_minimal_hgrid() +
  ylim(c(-0.65, 1.6)) +
  guides(color=FALSE)
```


```{r}
grid <- plot_grid(rapa_lines, oler_lines, labels = c('A', 'B'))

grid <-
plot_grid(
grid,
napus_lines,
labels = c('C', NULL),
label_y = 0,
nrow = 2
)
grid
```
```{r}
grid2 <-
  plot_grid(
  rapa_lines + xlab(NULL) + ylab('SHAP value'),
  oler_lines + ylab(NULL) + xlab(NULL),
  napus_lines + ylab(NULL) + xlab(NULL),
  nrow = 1,
  labels = c('A', 'B', 'C')
  )
grid2 <-
  ggdraw(
  add_sub(
  grid2,
  "Position on chromosome",
  vpadding = grid::unit(0, "lines"),
  y = 4.7,
  x = 0.5,
  vjust = 4.5
  )
  )
  
grid2
save_plot(grid2, filename = 'Shap_differences_Fig5.png', base_height = 6)
```



